


%\phantomsection
%\addcontentsline{toc}{chapter}{Preface}


%\begin{center}
\vspace{.25in}

This thesis signifies the completion of the \textit{Master of Science in Mathematical Sciences} program, representing the culmination of the author's four-year academic journey at NTNU.

I am grateful to my supervisor, \textsc{Sigrid Grepstad}, for her valuable guidance, detailed feedback, and generous allocation of time. Her expertise and insightful judgment have significantly contributed to the development of this thesis. I would also like to thank \textsc{Winston Heap}, my co-supervisor, for his helpful contributions, feedback, and suggestions during the final stages of completing this thesis; and \textsc{Jeffrey C. Lagarias} for both answering my queries and with the problem proposal for the \emph{American Mathematical Monthly}.   

I would also like to thank \textsc{Realfagskjelleren} — the student-driven bar I have been a part of, and \textsc{Linjeforeningen Delta} — the student association for my field of study, for providing a welcoming space of social gatherings, great friendships, and countless wonderful memories. 

I want to express my gratitude to my marvelous roommates of \emph{\enquote{the house with two rooms}} \textsc{Eg-51}, the individuals in \textsc{Reading Room 395b} for creating the finest environment for thesis writing, my e-sports team \textsc{N4F — Norges 4 Fronter} for conducting simulated counter-terrorism with gusto and bravery, and \textsc{all} the fellow students I have had the pleasure of getting to know during my time here in Trondheim. Your support and camaraderie have been truly appreciated. You are the bestest of the best, making Trondheim one of the best places to be. 

Lastly, I want to give a special mention to \textsc{Mathias N.}, \textsc{Erlend L.}, and \textsc{Håkon M.}, for (almost) always being available for coffee breaks; my \emph{algebros} and mathematics student partners \textsc{Elias A.}, \textsc{Talak M.}, \textsc{Preben S.}, \textsc{Nora A.}, and \textsc{Johan M.} for their support, discussions, and feedback; as well as \textsc{Mathias K.} and \textsc{[insert name]} for the final proofreading.

% I am grateful to my supervisor, Sigrid Grepstad, for her invaluable guidance and detailed feedback. Her generosity with her time and valuable insights have significantly contributed to the development of this thesis. I would also like to thank Winston Heap, my co-supervisor, for his helpful contributions, feedback, and suggestions during the final stages of completing this thesis.

%OPTIONAL: Before we commence the thesis, here are two quotes that sum up the thesis writing  
\vspace*{1cm}

\begin{flushright}
    Max Lunde Hauge\\
    Trondheim, Mai 2023
\end{flushright}

\clearpage

\newgeometry{top=0.5in, bottom=0.5in, left=1in, right=1in} % Adjust margins if needed
\thispagestyle{empty} % Remove page number

\vspace*{\fill}
\begin{center}
    \textit{\enquote{Oh, he seems like an okay person, except for being a little strange in some ways. All day he sits at his desk and scribbles, scribbles, scribbles. Then, at the end of the day, he takes the sheets of paper he's scribbled on, scrunches them all up, and throws them in the trash can.}}\\
\end{center}
\begin{flushright}
    \textbf{— J. von Neumann's housekeeper, describing her employer}
\end{flushright}
\vspace*{\fill}

\restoregeometry % Restore original page margins





%\end{center}